### 2021-09-07 README.Rmd
added readme skeleton

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# trolleyMultilabReplication

<!-- badges: start -->
<!-- badges: end -->

The goal of trolleyMultilabReplication is to provide

## Installation

The package is currently not on CRAN, but the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("marton-balazs-kovacs/trolleyMultilabReplication")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(trolleyMultilabReplication)
## basic example code
```